### 2025-12-05 Paper/paper_2025.tex
Updates from Overleaf

	\linespread{1.2}\title{\vspace{-0.5in} Overlapping Board Members in the Hospital Industry}

@@

    \begin{abstract}
		{\small
    Consolidation in health care has been a heavily discussed topic in research and regulation, as more concentration is linked to higher prices and lower quality. Aside from formal mergers and acquisitions, there is opportunity for hospitals to form affiliations with other competing hospitals through sharing common board members, which has mainly been studied in the context of large public firms. This paper establishes the existence of overlapping board members in competing nonprofit hospitals in the US. Using a data set constructed from IRS Tax Form 990s, I show that 8\% of this sample of nonprofit hospitals have overlapping board members with an otherwise unaffiliated hospital, and 17\% of hospital referral regions have overlapping boards within the market. Having an overlapping board member is correlated with quality metrics and investment decisions, though I do not draw a causal link. These findings highlight an understudied mechanism through which hospitals form affiliations, potentially affecting patient and market outcomes. 
		} 
	\end{abstract}

@@

    % Consolidation in health care remains a forefront concern for regulators in the US, as both horizontal and vertical integration have substantially increased over the last several decades (\cite{levinson2024ten}). Researchers have documented higher prices (\cite{brot2024pays}), and mixed effects on quality (\cite{kessler2000hospital}; \cite{cooper2011does}) as a result of health care becoming more concentrated through mergers and acquisitions. There is significant antitrust scrutiny in this sector, and proposed mergers are regularly denied due to expected harmful effects to competition (\cite{Meyer2022}). Firms in other industries have been shown to engage in methods of informal affiliation besides formal mergers and acquisitions. For example, instances of board overlap in public firms have been cited to violate the Clayton Act due to their anti-competitive nature. Yet, despite concerns of low competition, this type of connectedness among firms has not been given much consideration in health care. In this paper, I document the extent of overlapping board members in nonprofit hospitals in the US, and explore avenues of future research to investigate 
    % potential anti-competitive effects of this type of affiliation.

    % The presence of overlapping board members across firms, also referred to as board interlock, has been heavily studied in corporate finance, particularly for large publicly traded firms. Board overlap can facilitate indirect collusion on pricing and output strategies, and thus by the Clayton Act of 1914, board overlap among sufficiently large competing firms is illegal. Yet, it is known to exist even among the largest publicly traded firms. Researchers have found that board overlap is correlated with governance practices, performance, investment in research and development, and product differentiation (\cite{cai2014board}; \cite{lamb2016ties}; \cite{geng2021does}). However, it is still unclear whether board overlap really is anti-competitive, as it could also facilitate valuable knowledge sharing and allow multiple firms to gain from expertise of high quality board members. While the Department of Justice has increased scrutiny of board overlap, their focus has remained on tech and private firms (\cite{Morse2023}). Board overlap does occur in health care with instances of organizations denied merger permission due to boards not acting independently (\cite{huberfeld2006tackling}), of hospitals declaring informal affiliations through leadership (\cite{barnett_babcock_2012}), and of life sciences companies with substantial documented overlap (\cite{manjunath2024illegal}). However, little is known about how prevalent board overlap is in the hospital industry, and whether it is correlated with other hospital characteristics.

    % In this paper, I form networks of hospital board of director members using Tax Form 990s, which are yearly declarations of the financial state of nonprofits in the US that also include information about who is leading the firm. Extracting these identities and matching to other sources of hospital information, I show that approximately 8\% of the nonprofit hospitals in the sample are affiliated with another hospital in the same market through shared board members, even excluding system or network affiliations. Among hospitals that share board members, two thirds of the hospital pairs are between two general adult hospitals. That is, a majority of the shared board members serve on directly competing hospitals in the same geographic market. In addition, half of the hospital pairs are among one large and one small hospital, and half of the hospital pairs are between two hospitals that do not belong to a system.

    % Informed by established research on board overlap and hospital consolidation, I describe categories of hospital behaviors that may be affected by board overlap: price/quality, investment decisions, and differentiation. I construct various measures of behavior in these categories, and show how they are correlated with board overlap for small/large and system/independent hospitals. While the purpose of this paper is not to establish a causal relationship between board overlap and hospital behavior, there are key observable differences in the average quality and investment decisions of hospitals with and without board overlap that motivate future research on the topic.

    %  This paper contributes to our understanding of board overlap in a setting not previously considered and our understanding of hospital affiliations apart from formal mergers and acquisitions. The findings have implications for policymakers in assessing potential antitrust violations in certain nonprofit settings, and for future research in identifying the causal effects of board overlap in hospitals.

    \section{Related Literature}\label{sec:relatedlit}

    \subsection{Overlapping Board Members}\label{sec:boardoverlaplit}

    One of the first studies of board overlap argues that firms are not independent, self-sufficient entities as typically assumed in economics, but are actually heavily interconnected via overlapping board members (\cite{dooley1969interlocking}). Dooley frames board interlock as ``the root of many evils", and finds several characteristics to be correlated with the likelihood of sharing a board member with another corporation: size, financial relationships with other companies, and management control. This led to a number of studies focusing on how board interlock affects corporate governance, including whether board overlap stretches individual board members too thin (\cite{ferris2003too}; \cite{field2013busy}), contributes to increases in backdating stock options (\cite{bizjak2009option}), or led to convergence in governance strategy and behavior (\cite{bouwman2011corporate}; \cite{chiu2013board}; \cite{cai2014board}).

    The literature thus establishes a strong link between board interlock and governance practices. However, the relationship between board overlap and firm behavior, particularly its impact on competition, remains less understood. Theoretically, the direction of this effect is ambiguous. On one hand, board interlock is often viewed as an anti-competitive practice that enables implicit collusion. Alternatively, firms may seek highly experienced and connected individuals for their boards, leading to valuable knowledge sharing and enhanced decision-making. Most studies in this area have focused on the financial performance implications of board overlap, with firm behavior considered primarily as a mediating mechanism.

    Two main studies analyze of the effect of board overlap on financial performance. First, \citeauthor{baran2017director} (\citeyear{baran2017director}) leverages entry and exit from the S\&P 500, along with variation in the local supply of directors, and finds that well-connected board members contribute positively to firm value (\cite{baran2017director}). Similarly, leveraging policies in the US that remove barriers to board overlap, \citeauthor{geng2021does} (\citeyear{geng2021does}) finds that board overlap leads to higher profitability (\cite{geng2021does}). \citeauthor{schonlau2009board} (\citeyear{schonlau2009board}) and \citeauthor{cai2012board} (\citeyear{cai2012board}) analyze financial performance after mergers and acquisitions of publicly traded firms with connected boards, finding that firms with board overlap are more likely to be involved in acquisitions, and that specific types of board connections lead to better financial performance for the acquirer after the transaction (\cite{schonlau2009board}; \cite{cai2012board}).

    Beyond financial performance, several studies explore the relationship between board overlap and firm investment decisions as a mechanism for changes in financial performance. One paper examines whether interconnected firms are more likely to invest in the same information technology, and finds a positive association between board overlap and investment in IT among firms with active boards, indicating that timely communication is an important factor in whether board overlap matters (\cite{cheng2021social}). Finally, \citeauthor{geng2021does} (\citeyear{geng2021does}) finds that lower research and development expenditure and more product differentiation are key mechanisms through which overlapping firms improve profitability (\cite{geng2021does}).

    Despite these insights, significant gaps remain in our understanding of board overlap. Research has largely focused on governance or firm-level performance, while firm behavior and the broader market implications of board overlap remain underexplored. Additionally, most studies have been limited to large, publicly traded firms. This focus stems from the assumption that nonprofits engage in less anti-competitive behavior (\cite{baer2014clayton}; \cite{aai2013section7}), making board overlap in such organizations less concerning. However, this assumption is flawed, particularly in industries like health care, where nonprofit hospitals dominate and have been shown to engage in anti-competitive practices (\cite{hulver2023ftc}). Further research is needed to assess the competitive effects of board overlap across different market settings, particularly in nonprofit sectors where such dynamics may have substantial implications for industry competition and consumer welfare. I draw from this literature to inform potential outcomes of interest when considering board overlap in hospitals: financial performance, mergers and acquisitions, investment, and product differentiation. I discuss in Section \ref{sec:hospbehaviors} how these outcomes relate to hospitals specifically.

    \subsection{Common Ownership}\label{sec:commonownerlit}

    As consolidation in investment companies and in many industries has increased over time, so has the prevalence of common ownership among publicly traded firms. That is, firms within the same industry and market are increasingly owned by the same investment companies. While a separate occurrence and literature than that of board overlap, there are numerous parallels between the two concepts. Unlike publicly traded firms, nonprofits do not have shareholders that receive dividend payments. However, in some ways, a board functions similarly to shareholders: they influence the strategic direction of the organization through information sharing and voting on firm decisions without actively managing day-to-day operations. Further, the literature on common ownership in relation to market concentration and economic incentives is more developed than that of board overlap. Thus, I draw on this literature to inform outcomes that are potentially related to board overlap in the hospital industry.

    Whether common ownership affects competition is a widely debated theme in the economics literature. Theoretical models predict that common ownership will lead to lower equilibrium output and higher markups through firms placing non-zero weight on competitors' profit (\cite{rubinstein1983competitive}; \cite{rotemberg1984financial}; \cite{azar2012new}). Empirical evidence has grown tremendously since the seminal studies by \citeauthor{he2017product} (\citeyear{he2017product}) and \citeauthor{azar2018anticompetitive} (\citeyear{azar2018anticompetitive}), who found that common ownership leads to greater consolidation (acquisitions, joint ventures, strategic alliances), and that common ownership led to higher prices in the airline industry, respectively (\cite{he2017product}; \cite{azar2018anticompetitive}). However, these findings have been critiqued for how they measure common ownership and endogeneity concerns (\cite{kennedy2017competitive}; \cite{lewellen2021does}).

    A number of studies have since attempted to estimate the effect of common ownership on a range of outcomes including prices, research and development, merger activity, and patent durability (\cite{gerardi2023critical}). Many of which document a statistically significant relationship between common ownership and the relevant outcome. However, recent studies using alternative identification strategies find that common ownership can actually increase competition through lower insider trading profits, more product development, and higher investments (\cite{chen2023does}; \cite{kini2024common}). Additionally, recent literature has considered common ownership in the context of private firms with venture capitalists, where consolidated ownership is even more prevalent, and has found that competition decreases and information sharing increases in this setting (\cite{lindsey2008blurring}; \cite{gonzalez2020exchanges}; \cite{li2023common}; \cite{eldar2024common}).

    The literature on common ownership yields several conclusions for a study on board overlap. The first is a theoretical foundation confirming the importance of particular outcomes relevant to competition and consumer welfare, including merger and acquisition activity and investment decisions. Second, the common ownership literature focuses much more on how prices are impacted, which is also relevant in the health care setting. Finally, this literature introduces important concerns about the endogenous choice of owners or board members. While I do not seek to causally estimate the effects of board overlap in hospitals in this paper, this is an important consideration for future research.

    \subsection{Consolidation in Hospitals}\label{sec:hospconsol}

    There is a rich empirical literature examining the effects of consolidation in health care that includes both vertical and horizontal integration between hospitals, practices, and physicians. As the focus of this paper is affiliation among competing hospitals, I draw on the literature examining hospital consolidation in particular through mergers and systems. This will inform hospital behaviors that might be correlated with board overlap but are unique to the hospital setting and thus weren't discussed in Section \ref{sec:boardoverlaplit}. The main outcomes of interest in this area of research have centered around price and quality of care.

    Early studies on hospital consolidation found very little evidence of impacts on operating costs, efficiency, or quality of care (\cite{alexander1996short}; \cite{ho2000hospital}; \cite{dranove2003hospital}). From 2000-2020, the occurrence of hospital mergers increased dramatically, where a quarter of hospital markets included zero independent hospitals by 2020 (\cite{ElevanceHealth2023}). This increase in concentration has led to a vast number of studies estimating its effect on quality, operating costs, and prices.

    Studies typically document cost and efficiency gains experienced by hospitals who merge to become part of a larger system (\cite{schmitt2017hospital}; \cite{andreyeva2024corporatization}), even if the effects are small (\cite{craig2021mergers}). Naturally, this leads to the question of whether such efficiency gains are passed through to patients by lower prices. Across various settings, data sources, and methodology, hospital mergers are associated with significant increases in price due to hospitals having greater negotiating power with insurance companies (\cite{gaynor2012impact}; \cite{boozary2019association}; \cite{cooper2019price}; \cite{andreyeva2024corporatization}).

    There is mixed evidence on whether such gains in efficiency translate to better quality of care.  Several studies comparing hospitals that became a part of a system with observably similar hospitals that did not merge find little to no effect on readmission or mortality rates on average (\cite{haas2011mergers}; \cite{beaulieu2020changes}). However, one study finds that among heart disease patients, treatment intensity and mortality rates both increase after hospital consolidation (\cite{hayford2012impact}), and another finds increased readmission rates across all types of patients (\cite{andreyeva2024corporatization}).

    A select number of studies investigate other dimensions that could be affected by more system affiliation over time. In a slightly different setting of acquisitions of dialysis centers, \citeauthor{eliason2020acquisitions} (\citeyear{eliason2020acquisitions}) finds that decreases in quality are driven primarily by the convergence of treatment styles among acquired centers and their acquirer (\cite{eliason2020acquisitions}). Additional studies have suggested that consolidated service offerings is a potential mechanism for effects on quality (\cite{mariani2022impact}). Finally, \citeauthor{desai2023hospital} (\citeyear{desai2023hospital}) examine admission patterns of low-income patients as a result of increased hospital concentration, and find that Medicaid admissions decrease as a result of increased concentration. That is, highly concentrated markets exhibit redistribution patterns of patients within the market (\cite{desai2023hospital}).

    This literature informs ways that hospital affiliations might affect behaviors and outcomes specific to hospitals. In addition to the behaviors discussed in Sections \ref{sec:boardoverlaplit} and \ref{sec:commonownerlit}, measures of quality of care, treatment styles, service offerings, and admission patterns could also be important aspects driven by affiliation with other hospitals. I discuss these in more detail in Section \ref{sec:hospbehaviors}.

    \subsection{Informing Hospital Behaviors}\label{sec:hospbehaviors}

    The main purpose of this paper is to document the prevalence of board overlap in US hospitals in recent years. While I do not establish any causal relationship between board overlap and hospital behaviors, I consider the literature on board overlap, common ownership, and hospital consolidation discussed in Sections \ref{sec:boardoverlaplit}-\ref{sec:hospconsol} to inform a set of hospital behaviors that may be relevant for future research. I divide these behaviors into three categories: price/quality, investment decisions, and differentiation.

    The hospital consolidation literature highlights price and quality of care as key factors in the industry. Mergers often grant hospitals greater market power, enabling them to negotiate higher prices with insurance companies. It is unclear whether shared board members would lead to similar price increases. If board overlap facilitates information sharing about negotiations, it could indirectly influence pricing. However, direct board involvement in pricing decisions seems less likely. Nevertheless, with access to data on actual prices charged by hospitals, this remains an interesting outcome for future research to explore.

    It may be more likely that quality of care could be impacted by overlapping board members among hospitals. There are direct actions that could lead to changes in quality such as convergence in treatment styles or investment in technology, which I will discuss in detail in the following sections. There is also an indirect path through which overlapping board members might impact quality, which is that hospitals might compete for the same pool of experienced board members. If the board members with the most expertise are in high demand, sharing that expertise among multiple hospitals could facilitate greater improvements in quality.

    In both board overlap and common ownership, several investment decisions are shown to be associated with established firm affiliation. First, hospitals with board shared board members might be more likely to merge in the future. Second, sharing board members could facilitate engagement in the same types of technology or capital. Therefore, an area of future research is to look at outcomes such as health information technology investments, specific medical technology advancements in the hospital, and capital investments.

    Finally, previous research has shown that product differentiation is associated with board overlap in other industries. That is, sharing information among competing firms can lead to more specialization. In hospitals, there are two distinct types of differentiation: either through services offered, or the types of patients admitted. Both of these behaviors are key findings in the hospital merger literature, indicating that affiliations through board members could also facilitate changes in these areas. In future research, one could think about measuring the likelihood of specialized services being offered, where hospitals might shift investments towards services not offered by their affiliated hospital. Additionally, patient composition in the hospital, measured by the payer mix of admitted patients or complexity of their conditions, could be compared among hospitals with and without shared board members.

@@

@@

   I compile data on nonprofit hospital board of directors in years 2017-2022 from publicly available Tax Form 990s, which all sufficiently large nonprofits must file with the Internal Revenue Service (IRS) each year. These forms contain a section in which nonprofits declare their board, executives, and highest compensated employees. This paper focuses only on the board of director identities. I limit to firms that fill out Schedule H of the tax form, an indicator that the firm operates a hospital. I extract all names and positions of individuals listed in the section titled ``Officers, Directors, Trustees, Key Employees, and Five Highest Compensated Employees". I also record information on the hospital that filed the tax form: the Employee Identification Number (EIN), organization name and location, and other characteristics. The initial extraction yields 2,096 nonprofit hospitals.

   Next, I match hospitals in the tax form data to hospitals in other publicly available data sources. The only common information about hospitals across these sources is the name and location. Therefore, I use fuzzy string matching methods to create a crosswalk of EINs to identification numbers in the American Hospital Association (AHA) survey data. Specifically, I compute the Jaro-Winkler distance between hospital names in each data set. Then, I record matches with sufficiently similar names and addresses, yielding a sample of 1,836 matched hospitals.

    I limit the individuals in the sample to those on the board of directors, excluding executives that also serve on the board. I then record self-reported information about the board members in the data, including whether they include a title for medical doctor, nurse, or having a degree in health administration. I also record whether the firm records that they are a non-voting member of the board (though this is very rare). Several steps are necessary to ensure the names are clean and consistent across firms to identify people across hospitals. First, I remove any titles or credentials, and I convert common nicknames to their longer versions for consistency. Second, I combine names within the same firm that have either slight misspelling differences or, if middle names are only sometimes listed, they have at least 2 names in common. For example, if John Lee Matthew is a board member of hospital A in 2017, and John Matthew is a board member of hospital A in 2018, I assume these names represent the same person. I use the gender package in R to predict each individual's gender using historical data (\cite{gender}).

    \import{Objects}{board_people.tex}

>>>>

	\linespread{1.2}\title{\vspace{-0.5in} Unobserved Firm Affiliations: Evidence from Hospital Board Overlap}

@@

    \begin{abstract}
		{\small

		} 
	\end{abstract}

@@

    Antitrust enforcement in the US has been particularly relevant in recent years as price gouging and consolidation has increased.

    Antitrust legislation creates an additional hurdle for large-scale anti-competitive behavior, but there are other types of firm affiliations that typically elude antitrust scrutiny such as common ownership, joint ventures, and other types of affiliation/partnerships. By and large, firms form partnerships with other firms, often competitors, without needing to seek approval. This has started to raise concern recently, and states have even begun regulating smaller-scale affiliations in recent years. Horizontal mergers and consolidation have rightfully received a lot of attention in the literature, often linked to something?? However, we know little about the effects of smaller-scale affiliations, which could carry some of the same implications as consolidation proposals scrutinized for their ant-competitive nature.

    In this paper, I examine the effects of soft consolidation in the form of overlapping board members in nonprofit hospitals. Existing research establishes a strong link between board overlap and governance practices, but we know little about whether there should be anti-competitive concerns when firms share common board members. Additionally, most studies have been limited to large, publicly traded firms. This focus stems from the assumption that nonprofits engage in less anti-competitive behavior (\cite{baer2014clayton}; \cite{aai2013section7}), making board overlap in such organizations less concerning. However, this assumption is flawed, particularly in industries like health care, where nonprofit hospitals dominate and have been shown to engage in anti-competitive practices (\cite{hulver2023ftc}). Despite these insights, gaps persist: research rarely addresses market-level effects or nonprofit sectors, assuming nonprofits pose minimal antitrust risk (\cite{baer2014clayton}; \cite{aai2013section7}). This assumption is problematic in health care, where nonprofit hospitals dominate and exhibit anti-competitive behavior (\cite{hulver2023ftc}). Further work should explore competitive implications of board overlap beyond governance and financial metrics, particularly in nonprofit contexts.

    In 2024, the FTC banned a case of interlocking board (\cite{ftc2025sevita}).

@@

    \subsection{Nonprofit Hospital Governance}

    Nonprofits are governed at a high level by a board of directors: a group of unpaid community members that gather to oversee the broad direction and strategies of the firm. The board of directors does not mange the day-to-day activity of the hospital, but provides oversight to the highest level of manager, the Chief Executive Officer, whom the board also selects. The practical governance of boards is highly variable across different hospitals. State-level policies might influence details such as the regularity of board meetings or board member term limits. However, the main function of the board remains consistent: to provide strategic direction and oversight. In research focused on hospital governance, several factors have been found to be correlated with strategic change or hospital-level behaviors, including more influence over the CEO or top management (\cite{golden2001will}; \cite{alexander2008governance}; \cite{jiang2012enhancing}). Additionally, since the Affordable Care Act, boards have been actively involved in quality oversight (\cite{jha2010hospital}; \cite{prybil2014board}; \cite{prybil2010board}). Therefore, while boards are not usually involved in the day-to-day operations of hospitals, they do influence strategy and long-run hospital behavior.

    The board of directors selects a Chief Executive Officer, who then selects other members of the executive team such as a Chief Financial Officer and/or Chief Medical Officer. This leadership team manages the day-to-day operations of the hospital. An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize these teams. Some hospital executives specialize in health care administration by earning a degree in health care management, or an MBA specific to health care.

    In this paper, I consider the effects of board and executive interlock, someone sitting as an executive or on the board of multiple organizations. Board overlap does occur in health care with instances of organizations denied merger permission due to boards not acting independently (\cite{huberfeld2006tackling}), of hospitals declaring informal affiliations through leadership (\cite{barnett_babcock_2012}), and of life sciences companies with substantial documented overlap (\cite{manjunath2024illegal}). For many years, oversight of this activity in nonprofits was considered second-order to that of for-profits. However, in 2024, the FTC raised concern over board interlock of two prominent health care organizations (\cite{ftc2025sevita}). Additionally, board will sometimes ask executives from other organizations to sit on their board (\cite{gordon2025ftc}; \cite{alston2024governance}).

    Early research on board interlock highlights that firms are not isolated entities but interconnected through shared directors (\cite{dooley1969interlocking}), with interlocks linked to firm size, financial ties, and management control. Research has examined governance implications, including whether overlapping boards overextend directors (\cite{ferris2003too}; \cite{field2013busy}), increase stock option backdating (\cite{bizjak2009option}), or promote convergence in governance strategies (\cite{bouwman2011corporate}; \cite{chiu2013board}; \cite{cai2014board}). Empirical evidence focuses largely on financial outcomes: well-connected directors can enhance firm value (\cite{baran2017director}), and board overlap is associated with higher profitability (\cite{geng2021does}), greater acquisition activity, and improved post-merger performance (\cite{schonlau2009board}; \cite{cai2012board}). Mechanisms include shared IT investments (\cite{cheng2021social}), reduced R\&D, and increased product differentiation (\cite{geng2021does}).

    \subsection{Hospital Affiliations}\label{sec:hospaffil}

    Board interlock is often happening in conjunction with affiliation agreements: any relationship between otherwise independent health care providers designed to create shared advantage and value (\cite{aha2018affiliations}). Some examples of affiliation agreements include joint ventures for service provision, clinical agreements among hired health care workers, or joint management. It is common for these types of agreements to involve some sort of shared governance (\cite{aha2018affiliations}; \cite{natlawreview2023ftc}). While some have raised anti-trust concerns over affiliation agreements, they are largely undocumented, and therefore their effects are unknown (\cite{rand2023consolidation}). Organizations do not require pre-authorization to form affiliation agreements, but there are several states that now require notice of any type of affiliation (\cite{ncsl2024healthcare}).

    The more documented and studied type of hospital affiliation is that of traditional mergers and acquisitions. I will discuss the literature surrounding this topic in order to inform relevant behaviors that could also be affected by affiliation agreements. Hospital consolidation has accelerated over the past two decades, with mergers becoming so prevalent that by 2020, a quarter of hospital markets had no independent hospitals (\cite{ElevanceHealth2023}). Early studies found little evidence that consolidation improved operating costs, efficiency, or quality (\cite{alexander1996short}; \cite{ho2000hospital}; \cite{dranove2003hospital}), but more recent work documents modest efficiency gains for hospitals joining larger systems (\cite{schmitt2017hospital}; \cite{andreyeva2024corporatization}; \cite{craig2021mergers}). These gains, however, rarely translate into lower prices; instead, mergers consistently lead to significant price increases due to enhanced bargaining power with insurers (\cite{gaynor2012impact}; \cite{boozary2019association}; \cite{cooper2019price}; \cite{andreyeva2024corporatization}). Evidence on quality is mixed: most studies find little change in readmission or mortality rates (\cite{haas2011mergers}; \cite{beaulieu2020changes}), though some report higher treatment intensity and mortality for heart disease patients (\cite{hayford2012impact}) and increased readmissions overall (\cite{andreyeva2024corporatization}). Other research highlights mechanisms such as convergence in treatment styles (\cite{eliason2020acquisitions}), changes in service offerings (\cite{mariani2022impact}), and shifts in admission patterns, including reduced Medicaid admissions in concentrated markets (\cite{desai2023hospital}). These findings underscore that consolidation can influence not only costs and prices but also care delivery and patient access.

@@

   To create a measure of board/executive interlock, I compile data on nonprofit hospital board of directors and executives in years 2017-2022 from publicly available Tax Form 990s. All sufficiently large nonprofits must file these forms with the Internal Revenue Service (IRS) each year. They contain a section in which nonprofits declare their board, executives, and highest compensated employees. This paper focuses only on the board of director identities. I limit to firms that fill out Schedule H of the tax form, an indicator that the firm operates a hospital. I extract all names and positions of individuals listed in the section titled ``Officers, Directors, Trustees, Key Employees, and Five Highest Compensated Employees". I also record information on the hospital that filed the tax form: the Employee Identification Number (EIN), organization name and location, and other characteristics. The initial extraction yields 2,096 nonprofit hospitals.

   Next, I match hospitals in the tax form data to hospitals in other publicly available data sources. The only common information about hospitals across these sources is the name and location. Therefore, I use fuzzy string matching methods to create a crosswalk of EINs to identification numbers in the American Hospital Association (AHA) survey data. Specifically, I compute the Jaro-Winkler distance between hospital names in each data set. Then, I record matches with sufficiently similar names and addresses, yielding a sample of 1,824 matched hospitals.

     I record self-reported information about the individuals in the data, including whether they have a title for medical doctor, nurse, or having a degree in health administration. To clean the names, I remove any titles or credentials, and I convert common nicknames to their longer versions for consistency. I also combine names within the same firm that have either slight misspelling differences or, if middle names are only sometimes listed, they have at least 2 names in common. For example, if John Lee Matthew is a board member of hospital A in 2017, and John Matthew is a board member of hospital A in 2018, I assume these names represent the same person. Finally, I use the gender package in R to predict each individual's gender using historical data (\cite{gender}).

    \import{Objects}{boardand_people.tex}
